Cellular Automata and Computational Universality

### 2018-08-20 paper/paper.tex
Updates to paper. Started von Neumann's CA

    \begin{itemize}
        \item Class One: rapidly converge to a uniform state e.g. Rule 0, 232
        \item Class Two: rapidly converge to a repetitive or stable state e.g.
            Rule 250
        \item Class Three: appear to remain in a random state e.g. Rule 150, 182
        \item Class Four: form areas of repetitive or stable states, but also
            structures which interact in complicated ways e.g. Rule 54
    \end{itemize}

@@

    \subsection{Two dimensional cellular automata}
    Two dimensional cellular automata become a bit more interesting than their
    dimensionally-deficient brethren. Rather than having a single row of cells
    to consider at each generation, now a two-dimensional plane of cells is used
    to compute the evolution of the automaton. Whereas elementary automata
    looked only at the neighbourhood of $a_{i-1}, a_i,$ and $a_{i+1}$ when
    applying the ruleset, two-dimensional automata have a wider choice. Two of
    the most common types of neighbourhood to consider are the Moore
    neighbourhood and the von Neumann neighbourhood. The former includes the 8
    cells directly surrounding the current cell being considered, while the
    latter only focuses on the four orthogonally ajacent cells. Ocassionally,
    one may use the extended von Neumann neighbourhood.

    \subsection{Life-like cellular automata}
    \subsection{Cyclic cellular automata}

\section{Examples}
\section{Applications}
\section{Computational Universality}
\section{Firing Squad Synchronisation Problem}
\section{Majority Problem}

>>>>

    \begin{itemize}
        \item Class One: rapidly converge to a uniform state e.g. Rule 0, 232
        \item Class Two: rapidly converge to a stable or oscillating state e.g.
            Rule 250
        \item Class Three: appear to evolve in a chaotic fashion e.g. Rule 30,
            150, 182
        \item Class Four: form areas of repetitive or stable states, but also
            structures which interact in complicated ways e.g. Rule 54
    \end{itemize}

@@

    \subsection{Two dimensional cellular automata}
    Two dimensional cellular automata become a bit more interesting than their
    dimensionally-deficient brethren. Rather than having a single row of cells
    to consider at each generation, now a two-dimensional plane of cells is used
    to compute the evolution of the automaton. Whereas elementary automata
    looked only at the neighbourhood of $a_{i-1}, a_i,$ and $a_{i+1}$ when
    applying the set of rules, two-dimensional automata typically have a larger
    neighbourhood to consider. Two of the most common types of neighbourhood are
    the Moore neighbourhood and the von Neumann neighbourhood
    \cite{neighbourhoods}. The former includes the 8 cells directly surrounding
    the current cell being considered, while the latter only focuses on the four
    orthogonally ajacent cells. Ocassionally, one may use the extended von
    Neumann neighbourhood, for which the next four closest orthogonally adjacent
    cells are also included.

    Using a Moore neighbourhood and a set of states of size $k$, we can see that
    the total number of configurations for one particular neighbourhood is
    $k^9$, as each of the nine cells included in the neighbourhood may be in one
    of $k$ states. For an automaton such as Conway's Game of Life, where each
    cell is either alive or dead and a moore neighbourhood is used, there are
    $2^9=512$ ways to configure one of these neighbourhoods alone. The general
    equation for the number of automata for a given neighbourhood and number of
    possible states is $k^{k^n}$, where $k$ is the number of possible states and
    $n$ is the size of the neighbourhood used when determining the automaton's
    evolution between time steps (including the cell being computed); in the
    case of a two-dimensional automaton using two states and a Moore
    neighbourhood (such as Game of Life), there are $2^{2^9}$ possible automata
    \cite{num_rules}.

    \subsection{Game of Life}
    At this point it makes sense to talk a bit more about Conway's Game of Life.
    It was invented in 1970 by the British mathematician John Conway, after
    developing an interest in them after their introduction by John von Neumann
    in the late 1940s, who aimed to discover a hypothetical machine that was
    capable of recreating itself. Von Neumann was able to achieve this goal
    using a complex automaton, with 29 states and complicated rules to do so
    \cite{self-reproducing}. Von Neumann's automaton is discussed in more depth
    later. Conway's Game of Life was thus Conway's way of simplifying von
    Neumann's ideas, and in fact, the Game of Life is able to simulate the Game
    of Life itself and is in fact capable of universal computation.

    \subsection{Life-like cellular automata}

    \subsection{Cyclic cellular automata}

\section{Examples}
    \subsection{Von Neumann's automaton}
    \subsection{Game of Life}
    \subsection{Langton's Ant}
    \subsection{WireWorld}
\section{Applications}
\section{Computational Universality}
\section{Firing Squad Synchronisation Problem}
\section{Majority Problem}
